Towards Systematic Specification and Verification of Fault-Tolerant Protocols

### 2016-06-09 doc/om-paper/paper.tex
Add paper.bib and one citation

\newtheorem{definition}{Definition}

@@

\paragraph{Contributions}

Our work combines a number of ideas in the formal verification literature to build scalable, generic formal models of fault-tolerant distributed systems.

We describe a generic formal model for fault-tolerant distributed systems that aims to reduce the need for ad-hoc verification abstractions and optimizations while maintaining scalability in Section~\ref{sec:model}. The model combines several formalisms from the literature to efficiently but systematically model and verify a broad class of systems. The model extends \emph{calendar automata}, originally developed by Dutetre and Sorea~\cite{}. A calendar automata is a model inspired by discrete-event simulation, used in testing real-time systems~\cite{}. Calendar automata have previously been used to verify bounded asynchronous systems~\cite{}; here we verify a synchronous system.

@@

An \emph{event calendar} $\{ (e_0, t_0), (e_1, t_1), \ldots, (e_n, t_n) \}$ is a finite set of ordered pairs $(e_i, t_i)$ called \emph{calendar events} where $e_i \in E$ is an event and $t_i \in \mathcal{R}$ is a \emph{timeout}, the time at which the event is scheduled. We denote element $(e_1, t_1)$ of an event calendar by $c_i$.

@@

Let a transition system $\mathcal{M} = (S, I, \rightarrow)$, be a set of states $S$, a set of initial states $I \subseteq S$, and a transition relation $\rightarrow \subseteq S \times S$. We implicitly assume a set of state variables such that a state $\sigma \in S$ is a total function that maps state variables to values. We distinguish two special state variables in a transition system: (1) $now \in \mathcal{R}$ denotes the current time in the state, and (2) $c$ is an event calendar.

@@

\begin{itemize}
\item Time is initialized to be less than or equal to every calendar timeout: $\forall \sigma \in I, \, \forall (e_i, t_i) \in \sigma(c)$, $\sigma(now) \leq t_i$.
\item In all states, every calendar event occurs no sooner than the current time: $\forall \sigma \in S \, \forall (e_i, t_i) \in \sigma(c)$, $t_i \geq \sigma(now)$.
\item In all states, if the current time is strictly less than every calendar event, then the only enabled transition is a \emph{time progress} update: $\forall \sigma \in S \, \forall (e_i, t_i) \in \sigma(c)$, if $\sigma(now) < t_i$, then $\forall \sigma'$ such that $\sigma \rightarrow \sigma'$, $\sigma' = \sigma[now := min(c)]$.
\end{itemize}

\emph{give theorems for calendar, describe calendar for messages---all you need}

\subsection{Synchronous Observers}\label{sec:sync}

@@

\subsection{Abstract State Machines}\label{sec:abstract}

@@

\subsection{Byzantine Agreement}\label{sec:byz}

@@

>>>>

\newtheorem{definition}{Definition}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}

@@

Our contributions combine a number of ideas in the formal verification literature to build scalable, generic formal models of fault-tolerant distributed systems.

We describe a generic formal model for fault-tolerant distributed systems that aims to reduce the need for ad-hoc verification abstractions and optimizations while maintaining scalability in Section~\ref{sec:model}. The model combines several formalisms from the literature to efficiently but systematically model and verify a broad class of systems. The model extends \emph{calendar automata}, originally developed by Dutertre and Sorea~\cite{Dutertre-Sorea}. A calendar automata is a model inspired by discrete-event simulation, used in testing real-time systems~\cite{}.  Calendar automata have previously been used to verify bounded asynchronous systems~\cite{}; here we verify a synchronous system.

@@

An \emph{event calendar} $\{ (e_0, t_0), (e_1, t_1), \ldots, (e_n, t_n) \}$ is a set of ordered pairs $(e_i, t_i)$ called \emph{calendar events} where $e_i \in E$ is an event and $t_i \in \mathbb{R}$ is a \emph{timeout}, the time at which the event is scheduled. We denote element $(e_1, t_1)$ of an event calendar by $c_i$.

@@

Let a transition system $\mathcal{M} = (S, I, \rightarrow)$, be a set of states $S$, a set of initial states $I \subseteq S$, and a transition relation $\rightarrow \subseteq S \times S$. We implicitly assume a set of state variables such that a state $\sigma \in S$ is a total function that maps state variables to values. We distinguish two special state variables in a transition system: (1) $now \in \mathbb{R}$ denotes the current time in the state, and (2) $c$ is an event calendar.

@@

\begin{enumerate}
\item Time is initialized to be less than or equal to every calendar timeout: $\forall \sigma \in I, \, \forall (e_i, t_i) \in \sigma(c)$, $\sigma(now) \leq t_i$.
\item In all states, every calendar event occurs no sooner than the current time: $\forall \sigma \in S \, \forall (e_i, t_i) \in \sigma(c)$, $t_i \geq \sigma(now)$.
\item In all states, if the current time is strictly less than every calendar event, then the only enabled transition is a \emph{time progress} update: $\forall \sigma \in S, \, \forall (e_i, t_i) \in \sigma(c)$, if $\sigma(now) < t_i$, then $\forall \sigma'$ such that $\sigma \rightarrow \sigma'$, $\sigma' = \sigma[now := min(c)]$.
\item In all states, if the current time equals a timeout, then the only transition enabled is a calendar event update associated with the timeout: $\forall \sigma \in S, \, \exists (e_i, t_i) \in \sigma(c)$ such that $\sigma(now) = t_i$, then $\forall \sigma'$ such that $\sigma \rightarrow \sigma'$, $\sigma'(now) = \sigma(now)$, $\sigma'(c_j) = \sigma(c_j)$ for all $c_j \in \sigma(c)$ such that $c_j \neq c_i$, and $\sigma'(t_i) > \sigma(t_i)$.
\end{enumerate}

\begin{lemma}[Monotonic time]
$\forall \sigma, \sigma' \in S$, if $\sigma \rightarrow \sigma'$, then $\sigma'(now) \geq \sigma(now)$.
\end{lemma}

In a distributed system, it is convenient to distinguish global actions and local actions. Global actions are principally interprocess communication, while local actions are those carried out by each process to update its local state and produce new messages to broadcast. While both global and local actions can both be modeled as events in a calendar automata, doing so is generally overkill and complicates the model. Real-time constraints matter at the system-level but rarely within an individual process. From the global perspective, individual processes can update their local state atomically.

Following Dutetre and Sorea, we associate calendar events with channels in a distributed system~\cite{dutetre}. Specializing calendars to message passing does not lose generality since all external communication from an individual process can be abstracted as message passing. Furthermore, typical fault models can be abstracted to act over channels rather than processes~\cite{abstractions}. The calendar introduces real-time constraints on when processes send and receive messages.

Assume processes are indexed with the natural numbers $\mathbb{N}$. A \emph{channel} from process $i$ to $j$ is an ordered pair $(i,j)$. Define a set of messages $Msg$, with a designated element $null \in Msg$ denoting the absence of a message. Given a channel and a timeout, let $send$ return the message sent on the channel at that time:
$$send : \mathbb{N} \times \mathbb{N} -> \mathbb{R} -> Msg$$
So $send(i, j, t)$ returns the message that $i$ sends to $j$ at time $t$. Likewise, let
$$recv : \mathbb{N} \times \mathbb{N} -> \mathbb{R} -> Msg$$
compute the message received on that channel at that time, so that $recv(i, j, t)$ returns the message received by $j$ from $i$ at time $t$.

We require that messages received were previously sent and not previously received: if $recv(i, j, t) \neq null$, then $\exists t'$ such that $send(i, j, t')$ where $t' < t$, $recv(i, j, t) = send(i, j, t')$, and $\neg\exists t''$ such that $t' < t'' < t$ and $recv(i, j, t'') = recv(i, j, t)$

Then an event calendar for sending and receiving messages on channels is a set of send or receive events together with their timeouts that yield non-null messages; that is,
$$
\{ (e, t) | \forall i,j: e \neq null \textnormal{ and } (e = send(i, j, t) \textnormal{ or } e = recv(i, j, t))
\}
$$

\lee{technically, atomic time is just a convenience.}
The event of receiving a message initiates a process to update its local state machine and generate additional messages to send. When the process is updating its local state machine, the event calendar is paused. That is, updating an event $recv(i, j, t)$ implicitly includes updating $j$'s state machine.

\lee{note that we don't need full generality of send and receive events in OM(1)}

\subsection{Synchronous Observers}\label{sec:sync}
Synchronous observers are state machines that are synchronously composed with a system specification to check properties over the system's state variables. Synchronous observers are a method for composing specifications with a program. Synchronous observers were first introduced in the context of the synchronous programming languages, such as Lustre~\cite{}. While synchronous observers have been part of model-checking folklore for some time, Rushby systematically describes their application to the domain~\cite{}. The general form for specifying synchronous observers is the following:
$$system | assumptions | requirements$$
\noindent
where `$|$' denotes synchronous composition of state machines.

@@

% ------------------------------------------------------------
\section{Abstract State Machines}\label{sec:abstract}

% ------------------------------------------------------------
\section{An Extended Example: Byzantine Agreement}\label{sec:byz}

@@

@@

\bibliographystyle{IEEEtran}
\bibliography{paper}
